From Big Data to Big Displays\\ High-Performance Visualization at Blue Brain

### 2017-05-29 paper.tex
Motivation

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

@@

High-fidelity display systems are the integration point of the Blue Brain
visualization infrastructure. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and collaboration. Compared to
the current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a simple to use environment for high-fidelity visualization.
For all use cases, the increased display size and resolution allows improved
data exploration for 2D and 3D content.

@@

In the last years we created a production-ready setup of monoscopic tiled
display walls. On one hand, while there is substantial research on tiled display
wall software, we found that most solutions where not ready to be used in
$24\times 7$ unattained environment. For this reason, we started with the TACC
DisplayCluster open source software~\cite{DisplayCluster}, which we
incrementally refactored and improved to the current TIDE implementation.

@@

\begin{wrapfigure}{r}{.382\textwidth}\center\vspace{-6ex}
  \includegraphics[width=.382\textwidth]{images/ZeroMS}
  {\caption{\label{fZero}Messaging libraries and use cases}}\vspace{-2ex}
\end{wrapfigure}

@@

The pub-sub service is stateless, that is, applications have no expectation of
when messages are received or who receives published messages. This
communication pattern naturally leads to robust services, since there is no
possiblity for deadlocks or undefined behaviour. The pub-sub API is provided in
two flavors: a simple $pointer \& size$ memory buffer, and a higher level
object-based abstraction. The object-based API is syntactic sugar for the
low-level API, and allows automatic publish and update of objects with a few
lines of code. It uses the \textsf{toBinary()} and \textsf{fromBinary()} methods
of the \textsf{Serializable} interface to call the low-level API.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroBufStatic}\\\vspace{4ex}
  \includegraphics[width=\columnwidth]{images/ZeroBufDynamic}
  \caption{\label{fZeruBuf}ZeroBuf Examples for static (top) and dynamic (bottom) sized objects}
\end{figure}

@@

\fig{fJupyter} shows an example session of using this Python API from a Jupyter
notebook, allocating and launching a Brayns instance, setting relevant
rendering parameters and retrieving an image.

@@

The rendering applications form the backbone of our ecosystem. They cover a
wide range of established rendering algorithms to serve a broad set of use
cases for visual debugging, scientific illustrations and outreach.

@@

>>>>

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor Zero-Buf}

@@

\section{Motivation}

The Blue Brain Project (BBP) uses simulation-based research to analyze and reverse
engineer cortical neuron circuits. The simulated models go beyond using detailed
models of individual neurons or large-scale network models of simplified
neurons, they model in the order of hundreds of thousands of detailed neurons in
a fully connected circuit. The project generates a multitude of data for the
model building and during the simulation of these models. This data ranges from
terabyte-sized image stacks for data extraction to detailed in-silico circuit
models of large geometric complexity and terabyte-size simulation reports.

Visualization supports the BBP along all parts of this project to understand and
debug model data, building and simulation algorithms as well as validating and
discovering new insight from the in-silico experiments. Our strategy to support
this mission is based on components linked through network protocols:
High-fidelity display systems to see more detail in complex data, a set of
standard rendering engines (rasterization, out-of-core volume rendering,
interactive raytracing), and decoupled, light-weight applications using remote
these components. In the following we will present these components along with a
few use cases.

@@

High-fidelity display systems are the integration point of the Blue Brain
visualization infrastructure. They are the evolution of existing visualization
systems, enabling higher resolution, immersion and team collaboration. Compared
to current single-user or single-presenter systems, collaborative display
systems allow real team work through a combination of size, resolution and user
friendly implementation. Compared to immersive visualization systems like the
CAVE, they provide a simple to use environment for high-fidelity visualization.
For all use cases, the increased display size and resolution allows improved
data exploration for 2D and 3D content, facilitating large data exploration.

@@

In the last years we implemented a production-ready setup of monoscopic tiled
display walls. On one hand, while there is substantial research on tiled display
wall software~\cite{Sage, Sage2, Omegalib, DK:11, JLMV:06, DisplayCluster}, we
found that most solutions where not ready to be used in a $24\times 7$
unattained environment. For this reason, we started with the TACC DisplayCluster
open source software~\cite{DisplayCluster}, which we incrementally refactored
and improved to the current TIDE implementation.

@@

\begin{wrapfigure}{r}{.382\textwidth}\center\vspace{-6ex}
  \includegraphics[width=.382\textwidth]{images/ZeroMS}
  {\caption{\label{fZero}Messaging enables new use cases}}\vspace{-2ex}
\end{wrapfigure}

@@

The pub-sub service is stateless, that is, applications have no expectation of
when messages are received or who receives published messages. This
communication pattern naturally leads to robust services, since there is no
possiblity for deadlocks or undefined behaviour. The pub-sub API is provided in
two flavors: a simple $pointer\ \&\ size$ memory buffer, and a higher level
object-based abstraction. The object-based API is syntactic sugar for the
low-level API, and allows automatic publish and update of objects with a few
lines of code. It uses the \textsf{toBinary()} and \textsf{fromBinary()} methods
of the \textsf{Serializable} interface to call the low-level API.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/ZeroBufStatic}\\\vspace{4ex}
  \includegraphics[width=\columnwidth]{images/ZeroBufDynamic}
  \caption{\label{fZeroBuf}ZeroBuf Examples for static (top) and dynamic
(bottom) sized objects}
\end{figure}

\fig{fZeroBuf} shows two simple ZeroBuf schemas together with example usage of
the generated code in C++ and their memory layout. The static example shows
nested ZeroBuf classes for the camera used in \lref{lPubSub}, and the dynamic
example shows how raw data access is used to prepare a JPEG image for
publishing.

@@

\fig{fJupyter} shows an example session of using this Python API from a Jupyter
notebook, allocating and launching a Brayns instance, setting relevant rendering
parameters and retrieving an image. Note that the whole notebook runs in a
light-weight VM with no GPU and interacts with a Brayns instance launched on a
bare-metal visualization cluster node.

@@

The rendering applications form the backbone of our ecosystem. They cover a
wide range of established rendering algorithms to serve a broad set of use
cases for visual debugging, scientific illustrations and communication.

@@

Messaging enables surprising new use cases

Tiled display walls are cheap and ready for new ways of working
